### 2018-07-05 paper/3-abstractions.tex
Clarify what c stands for in Task c k v
Add a ref to Docker containers

\noindent
% Here \hs{c} stands for \emph{constraint}, such as \hs{Applicative} or \hs{Monad}.
A \hs{Task} describes a single build task, while \hs{Tasks} associates a
\hs{Task} to every non-input key; input keys are associated with
\hs{Nothing}. The highly-abstracted type\footnote{Readers
familiar with \emph{lenses} or \emph{profunctor optics} might recognise a
familiar pattern. We discuss this in~\S\ref{sec-related-optics}.} \hs{Task} is
best introduced by an example. Consider this \Excel spreadsheet:

@@

But nor can the task description work for \emph{any} \hs{f}; most task
descriptions (e.g. \hs{sprsh1} in \S\ref{sec-task}) require that \hs{f}
satisfies certain properties, such as \hs{Applicative} or \hs{Monad}. That is
why \hs{Task} has the ``\hs{c}~\hs{f}~\hs{=>}'' constraint in its type,
expressing that \hs{f} can only be instantiated by types that satisfy the
constraint \hs{c}. So the type \hs{Task} emerges naturally, almost inevitably.
But now that it \emph{has} emerged, we find that constraints \hs{c} classify
task descriptions in a very interesting way:
\begin{itemize}
\item \hs{Task}~\hs{Applicative}. In \hs{sprsh1} we needed only \hs{Applicative}
  operations, expressing the fact that the dependencies between cells can be
  determined \emph{statically}; that is, by looking at the formulae, without
  ``computing'' them (see \S\ref{sec-deps}).
\item \hs{Task}~\hs{Monad}. As we shall see in \S\ref{sec-task-monad}, a monadic
  task allows \emph{dynamic} dependencies, in which a formula may depend on cell
  \cmd{C}, but \emph{which} cell \cmd{C} depends on the value of another cell
  \cmd{D}.
%TODO: Add a reference for Docker?
\item \hs{Task}~\hs{Functor} is somewhat degenerate: the task description cannot
  even use the application operator \hs{<*>}, which limits dependencies to a
  single linear chain, as e.g. in the Docker build system. It is interesting to
  note that, when run on such a task description, the \hs{busy} build system
  will build each key at most once, thus partially fulfilling the minimality
  requirement~\ref{def-minimal}. Alas, it still has no mechanism to decide which
  input keys changed since the previous build.
\item \hs{Task}~\hs{Alternative}, \hs{Task}~\hs{MonadPlus} and their
  variants can be used for describing tasks with a certain type of
  non-determinism, as discussed in~\S\ref{sec-non-determinism}.
\item \hs{Task}~\hs{(MonadState}~\hs{i)} will be used
  in~\S\ref{sec-implementations} to describe tasks that have read and write
  access to the persistently stored build information~\hs{i}.
\end{itemize}

>>>>

\noindent
Here \hs{c} stands for \emph{constraint}, such as \hs{Applicative} or \hs{Monad}.
A \hs{Task} describes a single build task, while \hs{Tasks} associates a
\hs{Task} to every non-input key; input keys are associated with
\hs{Nothing}. The highly-abstracted type\footnote{Readers
familiar with \emph{lenses} or \emph{profunctor optics} might recognise a
familiar pattern. We discuss this in~\S\ref{sec-related-optics}.} \hs{Task} is
best introduced by an example. Consider this \Excel spreadsheet:

@@

But nor can the task description work for \emph{any} \hs{f}; most task
descriptions (e.g. \hs{sprsh1} in \S\ref{sec-task}) require that \hs{f}
satisfies certain properties, such as \hs{Applicative} or \hs{Monad}. That is
why \hs{Task} has the ``\hs{c}~\hs{f}~\hs{=>}'' constraint in its type,
expressing that \hs{f} can only be instantiated by types that satisfy the
constraint \hs{c}. So the type \hs{Task} emerges naturally, almost inevitably.
But now that it \emph{has} emerged, we find that constraints \hs{c} classify
task descriptions in a very interesting way:
\begin{itemize}
\item \hs{Task}~\hs{Applicative}. In \hs{sprsh1} we needed only \hs{Applicative}
  operations, expressing the fact that the dependencies between cells can be
  determined \emph{statically}; that is, by looking at the formulae, without
  ``computing'' them (see \S\ref{sec-deps}).
\item \hs{Task}~\hs{Monad}. As we shall see in \S\ref{sec-task-monad}, a monadic
  task allows \emph{dynamic} dependencies, in which a formula may depend on cell
  \cmd{C}, but \emph{which} cell \cmd{C} depends on the value of another cell
  \cmd{D}.
%TODO: Add a reference for Docker?
\item \hs{Task}~\hs{Functor} is somewhat degenerate: the task description cannot
  even use the application operator \hs{<*>}, which limits dependencies to a
  linear chain, as e.g. in Docker containers~\cite{docker}. It is
  interesting to note that, when run on such a task description, the \hs{busy}
  build system will build each key at most once, thus partially fulfilling the
  minimality requirement~\ref{def-minimal}. Alas, it still has no mechanism to
  decide which input keys changed since the previous build.
\item \hs{Task}~\hs{Alternative}, \hs{Task}~\hs{MonadPlus} and their
  variants can be used for describing tasks with a certain type of
  non-determinism, as discussed in~\S\ref{sec-non-determinism}.
\item \hs{Task}~\hs{(MonadState}~\hs{i)} will be used
  in~\S\ref{sec-implementations} to describe tasks that have read and write
  access to the persistently stored build information~\hs{i}.
\end{itemize}
